---
title: "Scripted Reruns and Saved Specs"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Scripted Reruns and Saved Specs}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

This is **Chapter 9 of 10** in the beginner path.

Up to this point, we have focused on the beginner path. Now we move one step toward reproducibility and automation.

The key idea in this chapter is simple: a *spec* is just a saved set of instructions for a `moover` run.

## Why saved specs are useful

When people run a workflow interactively, they often remember the broad choices but forget the exact settings. A saved spec solves that problem. It records the file locations and settings in a form that can be reused.

That means a spec is useful for at least three reasons:

1. you can rerun the same workflow later
2. you can share the run instructions with a colleague
3. you can move from interactive work to scripted work without changing the underlying pipeline

## Create a spec in code

Here is a simple example of a spec for a training workflow.

```{r, eval = FALSE}
library(moover)

spec <- create_spec(
  workspace = list(root = "my_moover_workspace"),
  labels = list(
    tech_file = "tech.csv",
    path = "observations.csv"
  ),
  features = list(
    selection = "standard",
    standard_set = "manual5"
  ),
  optimise = list(enabled = FALSE)
)
```

Once you have the spec, you can rerun the same workflow directly.

```{r, eval = FALSE}
run_pipeline(spec, stage = "all")
```

## Use a saved JSON spec

`moover` also writes specs to JSON so that runs can be reproduced later without stepping through the wizard again.

```{r, eval = FALSE}
# Re-run a workflow from a previously saved JSON spec.
run_pipeline("runs/20260330_012345/spec/run_spec.json", stage = "all")
```

You do not need to hand-edit JSON to benefit from this. Most beginners can think of the JSON file as a saved record of the choices they already made.

## A good workflow to aim for

A common pattern is:

1. start with the wizard
2. let `moover` save the spec for you
3. rerun that same spec later in code when you want something repeatable

That is a nice bridge between beginner and advanced use. You do not have to jump straight into scripting on day one.

<div class="moover-next-steps">
<strong>Move through the tutorial</strong><br>
Previous chapter: <a href="optimise-a-model-for-deployment.html">Chapter 8. Optimise a Model for Deployment</a><br>
Next chapter: <a href="troubleshooting-and-glossary.html">Chapter 10. Troubleshooting and Glossary</a>
</div>
