---
title: "Set Up Your Workspace"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Set Up Your Workspace}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

This is **Chapter 2 of 10** in the beginner path.

A `moover` workspace is just an ordinary folder on your computer. You do not need a high-performance cluster, and you do not need an `.Rproj` file. What you do need is a consistent place for raw data, run outputs, and a small number of support files.

In this chapter, we'll do three things:

1. create a workspace
2. look at the folder structure
3. decide where your main input files belong

## Create the workspace

The `init_workspace()` function creates the standard folder structure. If the folders already exist, it leaves them in place.

```{r, eval = FALSE}
library(moover)

# Create a moover workspace in a folder of your choice.
# If the folder does not exist yet, moover will create it.
ws <- init_workspace("my_moover_workspace")

# Print the key folder paths that moover will use.
ws
```

That one step gives you the basic structure `moover` expects.

## What the folders are for

A beginner-friendly workspace is helpful because the same questions come up every time: where do the raw files go, where are the model outputs, and where did the package save this run?

A standard `moover` workspace looks like this:

```text
my_moover_workspace/
  data_raw/
  runs/
  _internal/
  tech.csv
  observations.csv
```

Here is what those folders mean in practice.

### `data_raw/`

This is where your raw accelerometer files go. They stay here unchanged. `moover` reads from this folder; it does not expect you to hand-edit the files during a run.

### `runs/`

Every time you run a workflow, `moover` creates a new run folder here. That makes each run traceable. You can come back later and see exactly what happened during that run.

### `_internal/`

This folder is for helper files that `moover` writes for its own use. Most beginners do not need to work directly in this folder, but it is part of keeping the workflow reproducible.

## Where the main input files go

For a training workflow, two extra files usually sit at the top level of the workspace:

- `tech.csv`
- `observations.csv`

The `tech.csv` file links animal ids to accelerometer ids when that mapping is needed. The `observations.csv` file holds the labelled behaviour periods you want the model to learn from.

A common beginner setup looks like this:

```text
my_moover_workspace/
  data_raw/
    animal_01_cquFormat.csv
    animal_02_cquFormat.csv
  runs/
  _internal/
  tech.csv
  observations.csv
```

## What a run folder looks like

When you start a run, `moover` creates a self-contained folder inside `runs/`. The exact name changes from run to run, but the structure is consistent.

```text
runs/
  20260330_012345/
    spec/
    results/
    models/
    plots/
    qc/
```

That structure is useful because it means you can answer practical questions later:

- Which settings did I use?
- Which model export belongs to this run?
- Where are the plots and metrics?

## A note on runtime

`moover` is designed to run locally on an ordinary computer. Larger datasets can still take time, and that is normal. The main goal of the package is not to pretend the work is instant; it is to make the work manageable and understandable for beginners.

<div class="moover-next-steps">
<strong>Move through the tutorial</strong><br>
Previous chapter: <a href="getting-started.html">Chapter 1. Start Here</a><br>
Next chapter: <a href="accelerometer-input-formats.html">Chapter 3. Prepare Your Accelerometer Files</a>
</div>
