---
title: "Train and Predict with Real Data"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Train and Predict with Real Data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

This page covers **Chapter 5 and Chapter 6** of the beginner path.

This chapter holds **two walkthroughs on one page** because they are really two starting points for the same `moover` workflow.

- **Walkthrough 1** starts from labelled raw data and takes us all the way through training, validation, export, and prediction on a larger dataset.
- **Walkthrough 2** starts from a ready-made model bundle and shows how to apply it straight to raw data.

If you are learning `moover` as a full workflow, begin with Walkthrough 1. If you already have a model and only care about prediction, jump straight to Walkthrough 2.

The promise of `moover` is that this should stay manageable. Once your accelerometer files, tech details, and observations are in the expected formats, the package takes over the harder pipeline steps: feature calculation across fixed epochs, Random Forest training, leave-one-animal-out validation, export, and prediction on larger datasets.

## Walkthrough 1: Train, Validate, Export, Predict {#walkthrough-1}

In this first walkthrough, we will:

1. copy a packaged real-data training workspace into a folder we can write to
2. train and validate a Random Forest model
3. export that model as a reusable bundle
4. apply the newly exported model to a second, larger packaged raw dataset

### What we are using

This training example is a curated subset of real `RF_Embed_v2` data. The package ships it as a normal `moover` workspace so you do not have to assemble the files yourself.

```{r, eval = FALSE}
library(moover)

# This packaged folder contains:
# - data_raw/ with curated accelerometer files
# - tech.csv linking animal ids to accelerometer ids
# - observations.csv in moover's default labelled-bout schema
train_example <- system.file("extdata", "example_train_workspace", package = "moover")

# This second packaged folder contains a larger raw-only dataset.
# We will use it after training to show prediction beyond the labelled subset.
predict_example <- system.file("extdata", "example_predict_raw", package = "moover")
```

### Copy the training workspace to your computer

We copy the packaged training workspace before running anything because installed package files should be treated as read-only.

```{r, eval = FALSE}
# Choose a writable folder on your own computer.
workspace_dir <- file.path(tempdir(), "moover_real_data_walkthrough")
dir.create(workspace_dir, recursive = TRUE, showWarnings = FALSE)

# Copy the packaged training workspace so this run can write outputs into it.
file.copy(
  from = list.files(train_example, full.names = TRUE, all.files = TRUE, no.. = TRUE),
  to = workspace_dir,
  recursive = TRUE
)
```

### Create the run specification

This spec does two jobs.

First, it tells `moover` how to train the model from the labelled raw data in the copied workspace.  
Second, it tells `moover` to use the newly exported model to predict on the larger raw-only dataset after training finishes.

```{r, eval = FALSE}
# Create a spec that trains first and then predicts on a second raw dataset.
spec <- create_spec(
  workspace = list(root = workspace_dir),
  ingest = list(
    # Training raw data lives inside the copied example workspace.
    raw_dir = file.path(workspace_dir, "data_raw")
  ),
  labels = list(
    # These two files provide the training labels and the id mapping.
    tech_file = file.path(workspace_dir, "tech.csv"),
    path = file.path(workspace_dir, "observations.csv")
  ),
  features = list(
    # Keep the example simple and use the standard beginner feature set.
    selection = "standard",
    standard_set = "manual5"
  ),
  optimise = list(
    # We are teaching the main workflow here, so keep optimisation off.
    enabled = FALSE
  ),
  model = list(
    positive_class = "grazing",
    num_threads = 1L,
    num_trees = 3L,
    min_node_size = 30L
  ),
  predict = list(
    # After the model is exported, use it on the second packaged raw dataset.
    raw_dir = predict_example,
    predict_after_export = TRUE
  )
)
```

### Run the full workflow

This is the main beginner command. In this case, `stage = "all"` means:

1. import and check the training raw data
2. build 10-second epochs and calculate features
3. match those epochs to the labelled behaviour bouts
4. train and validate the model with LOCO
5. export the model bundle
6. use the new export bundle to predict on the larger raw dataset

```{r, eval = FALSE}
# Run the complete workflow from labelled raw data through to prediction.
run_pipeline(spec, stage = "all")
```

If you prefer prompts instead of code, the training wizard can guide you through the same path:

```{r, eval = FALSE}
# The training wizard can also ask whether you want to predict on another
# raw-data folder after export.
wizard_train()
```

### What success looks like

After the run finishes, there are three places to check.

1. **LOCO validation outputs**
   - these tell you how the model performed on animals it did not see during training
   - look for files such as `metrics_overall.csv`, `confusion_matrix.csv`, and `metrics_by_class.csv`
2. **The export bundle**
   - this lives in the run's `models/` folder
   - it contains the fitted model, feature manifest, tree dump JSON, metrics, and test vectors
3. **Prediction outputs**
   - these live in the run's `results/` folder
   - the main file is `epoch_predictions.csv`

This is the key idea of Walkthrough 1: we use a labelled subset to build the model, and then we immediately use that model on a larger raw dataset.

## Walkthrough 2: Predict With a Shipped Model {#walkthrough-2}

Now let's start from the other end.

In this walkthrough, we assume the model has already been built for you. You simply want to apply it to raw data.

We will:

1. locate the packaged example model bundle
2. point `moover` to the packaged prediction raw dataset
3. run prediction and inspect the output

### Locate the packaged model bundle

This bundle ships with the package and already contains the feature manifest and model settings needed for prediction.

```{r, eval = FALSE}
# This packaged folder contains the shipped example bundle.
# It includes rf_model_full.rds, feature_manifest.csv, export_config.json,
# model_spec.json, and rf_tree_dump.json.
bundle_dir <- system.file("extdata", "example_model_bundle", package = "moover")

# Reuse the packaged raw-only prediction dataset from Walkthrough 1.
predict_example <- system.file("extdata", "example_predict_raw", package = "moover")
```

### Create a prediction-only spec

Notice what is missing here:

- no observations file
- no feature list written by hand
- no epoch length chosen by hand

That is the point of the bundle. It tells `moover` what to rebuild.

```{r, eval = FALSE}
predict_workspace <- file.path(tempdir(), "moover_prediction_only")
init_workspace(predict_workspace)

predict_spec <- create_spec(
  workspace = list(root = predict_workspace),
  ingest = list(
    # Point to the raw accelerometer files you want to classify.
    raw_dir = predict_example,
    # This dataset is large enough that chunked reading is sensible.
    chunk_rows = 50000L
  ),
  labels = list(
    # No labelled observations are needed for prediction-only work.
    tech_file = "",
    path = NULL
  ),
  predict = list(
    # The model bundle contains the information needed to rebuild the features.
    model_bundle = bundle_dir
  )
)
```

### Run prediction

```{r, eval = FALSE}
# Predict behaviour on the packaged raw-only dataset.
run_pipeline(predict_spec, stage = "predict")
```

The prediction wizard can help with the same workflow if you prefer prompts:

```{r, eval = FALSE}
wizard_predict()
```

### What happens during prediction

During prediction, `moover`:

1. reads the raw accelerometer files
2. rebuilds the correct 10-second epochs
3. calculates the exact features required by the bundle
4. applies the model
5. writes `epoch_predictions.csv`

The important beginner message here is that the bundle carries the instructions. You do not manually restate the feature set or epoch definition when you already have a model bundle.

## When to use walkthrough 1 versus walkthrough 2

Use **Walkthrough 1** when:

- you have labelled observations
- you need to train or validate a model
- you want to export a new bundle for sharing or deployment

Use **Walkthrough 2** when:

- a model bundle already exists
- you mainly want predictions on new raw data
- you are acting as the collaborator applying a shared model rather than building one from scratch

<div class="moover-next-steps">
<strong>Move through the tutorial</strong><br>
Previous chapter: <a href="recording-observations.html">Chapter 4. Record Observations for Training</a><br>
Next chapter: <a href="deployment-export-format.html">Chapter 7. Understand Your Results and Export Folder</a>
</div>
